Exploring the Motives and Driving Forces Behind Modifications to APIs From a Software Evolution Perspective

### 2015-05-06 thesis/paper.tex
Notes

\subsection{RQ2 Results} \label{results_rq2}
\noindent
RQ2: \textbf{How do APIs ``passively'' evolve?}
\smallskip

@@

\section{Discussion}
Our results have provided valuable insight into the motives behind the evolution of APIs. The results show that APIs definitely can be studied from the point of view of long-established software evolution theories. Based on the results, we have formulated the subsequent outcomes.

\subsection{Discussion RQ1}

\noindent
\textbf{Functionality is a prime driver of API evolution  } 
Overwhelmingly, the primary driving factor behind API evolution was the desire for new API functionality. We found that a vast majority of the changes either added functionality or were a direct consequence of increased functionality. The main driver for these changes was the pressure from API users, but pressure from the case company developers who utilise the API also contributed. This focus upon functionality was also emphasised partially due to a business goal. The more features the API can provide, the more applications developers can create for the cameras. This has a domino effect of increasing the appeal and therefore value of the products.

\smallskip \noindent 
\textbf{Maintainability does not drive change  }  
While maintainability is important, it is not a primary driving factor for change. We found that where changes that effected maintainability, functionality had also been added or modified. Of the changes identified, none were considered to be primarily groomative. Groomative motives were only found to be secondary, and were never the main motive for a change. Changes to the example code are an instance of this. The example code was updated purely to reflect the changes in functionality and not as a groomative effort.

\smallskip \noindent
\textbf{Standardising application development  } 
Supporting users of the API in creating correct and robust applications is a very important evolution factor for the case company. Changes between the versions introduced improvements to error handling, the example code and the way in which the API utilises interfaces. This, coupled with removal of unwanted functionality and unsafe operations, has allowed the API to grow and become much more user-friendly. As mentioned previously, this could increase the amount of applications being created for the camera and as a consequence, could potentially increase the value of them.

\smallskip \noindent
\textbf{Removing functionality  } 
Previous literature has suggested that deprecating functionality is the most common technique used to remove support for specific functionality. Removals are only suggested in specific circumstances. The successful removal of the burst function seems to indicate that under the correct circumstances, removing specific functionality is an alternative that should be considered.
The case company removed this function from the API, after an internal review found that the potential negative effect the removal could have, would be less impactful than further maintenance of the function. We therefore recommend to remove functionality of an API if the effect of the removal has been thoroughly investigated and that the removals happen incrementally.

\smallskip \noindent
\textbf{Documentation does not drive change  } 
The intent to update example code and documentation was described by the interviewee to be an important change between version 1.4 and 2.0. While important for the APIs usability, it was not a driving factor of change itself. The changes to the APIs' design and functionality caused the need to update the documentation and example code.

\smallskip \noindent
\textbf{Hardware is not a driving factor  } 
Our initial expectations included the view that changes to hardware would be one of the strongest driving factors of API evolution. One example could be if new cameras are created or new hardware is added to an existing camera, which would require updates to the API to allow use of the new features. However, after analysing the changes and from the information gathered in the interview, we discovered that hardware has played a minimal role in the evolution of the API.

@@

\noindent
\textbf{Functionality is reverse engineerable  } 
All changes whose main motive was to either add or remove functionality were accurately predicted. These changes were the most significant and it was expressed during the interview that these types of changes were the main driving factors behind the development of the new version. In summary, eleven out of a total of thirteen changes were correctly predicted. It is also interesting to note that all changes which were not assigned any sub-category were correctly predicted. This supports the fact that such changes have a clearer intent and are perhaps more well-defined.

\smallskip \noindent
\textbf{Design changes are reverse engineerable to some extent} 
Three out of five changes which did not add functionality, but rather were introduced to reduce complexity and increase quality, were correctly predicted. The two changes which were incorrectly predicted were inaccurate because the main motivation was confused with the secondary intent. This shows that, although the prediction was incorrect, the general reasoning was correct. Our predictions were accurate enough to allow us to form a reliable conclusion which provided valuable discussion with the case company.

@@

\noindent
\textbf{A majority of the laws apply  }
Sufficient evidence was found for the laws of Continuing Change, Self Regulation, Conservation of Familiarity, Continuing Growth and Feedback System to establish that they apply to APIs. The most prominent explanation for this is that a large part of the introduced changes were related to functionality. It is especially interesting to note that the laws of Self Regulation and Conservation of Familiarity accurately match the API explored in this paper.

\smallskip \noindent
\textbf{Two laws apply to a certain extent  }
The laws of Increasing Complexity and Declining Quality were found to apply to APIs, but the extent to which they apply might not be as significant for APIs compared to other types of systems. The reason that this doubt exists is because these motives were secondary to the motives of increased functionality. An alternative explanation for these laws being of lesser significance in APIs could be that the general inertia to modify APIs which exists \cite{google_talk} \cite{henning2007api} \cite{mcdonnell2013empirical} \cite{robbes2012developers} might dominate the need to improve simplicity and quality.

\smallskip \noindent
\textbf{One law is not applicable  }
Strong evidence was found to suggest that the law of Conservation of Organisational Stability does not apply to APIs. Our results not only show that the law is not true for APIs, but that the opposite of what the law states is true for APIs. This is supported by the consensus in literature of the negative effects caused by modifications to APIs \cite{google_talk} \cite{henning2007api} \cite{mcdonnell2013empirical} \cite{robbes2012developers}. The results from our study also revealed that reluctance to change APIs causes the releases of new versions to be few and sporadic.

\smallskip \noindent
\textbf{The importance of a Feedback System  }
Especially interesting results were found in regards to the law of Feedback System. The importance of having a structured and developed feedback loop was stressed during the interview and it was also expressed as one of the main points of improvement. This suggests that including feedback from more sources and using this feedback to improve the evolution process should be a consideration for APIs.

>>>>

\subsection{RQ2 Results} \label{results_rq2} %Reviewed by JW
\noindent
RQ2: \textbf{How do APIs ``passively'' evolve?}
\smallskip

@@

\section{Discussion} %Reviewed by JW
Our results have provided valuable insight into the motives behind the evolution of APIs. The results show that APIs definitely can be studied from the point of view of long-established software evolution theories. Based on the results, we have formulated the subsequent outcomes.

\subsection{Discussion RQ1} %Reviewed by JW

\noindent
\textbf{Functionality is a prime driver of API evolution  }  %Reviewed by JW
Overwhelmingly, the primary driving factor behind API evolution was the desire for new API functionality. We found that a vast majority of the changes either added functionality or were a direct consequence of increased functionality. The main driver for these changes was the pressure from API users, but pressure from the case company developers who utilise the API also contributed. This focus upon functionality was also emphasised partially due to a business goal. The more features the API can provide, the more applications developers can create for the cameras. This has a domino effect of increasing the appeal and therefore value of the products.

\smallskip \noindent 
\textbf{Maintainability does not drive change  }  %Reviewed by JW
While maintainability is important, it is not a primary driving factor for change. We found that where changes that effected maintainability, functionality had also been added or modified. Of the changes identified, none were considered to be primarily groomative. Groomative motives were only found to be secondary, and were never the main motive for a change. Changes to the example code are an instance of this. The example code was updated purely to reflect the changes in functionality and not as a groomative effort.

\smallskip \noindent
\textbf{Standardising application development  } %Reviewed by JW%Reviewed by JW
Supporting users of the API in creating correct and robust applications is a very important evolution factor for the case company. Changes between the versions introduced improvements to error handling, the example code and the way in which the API utilises interfaces. This, coupled with removal of unwanted functionality and unsafe operations, has allowed the API to grow and become much more user-friendly. As mentioned previously, this could increase the amount of applications being created for the camera and as a consequence, could potentially increase the value of them.

\smallskip \noindent
\textbf{Removing functionality  } %Reviewed by JW
Previous literature has suggested that deprecating functionality is the most common technique used to remove support for specific functionality. Removals are only suggested in specific circumstances. The successful removal of the burst function seems to indicate that under the correct circumstances, removing specific functionality is an alternative that should be considered. The case company removed this function from the API, after an internal review found that the potential negative effect the removal could have, would be less impactful than further maintenance of the function. We therefore recommend to remove functionality of an API if the effect of the removal has been thoroughly investigated and that the removals happen incrementally.

\smallskip \noindent
\textbf{Documentation does not drive change  } %Reviewed by JW
The intent to update example code and documentation was described by the interviewee to be an important change between version 1.4 and 2.0. While important for the APIs usability, it was not a driving factor of change itself. The changes to the APIs' design and functionality caused the need to update the documentation and example code.

\smallskip \noindent
\textbf{Hardware is not a driving factor  } %Reviewed by JW
Our initial expectations included the view that changes to hardware would be one of the strongest driving factors of API evolution. One example could be if new cameras are created or new hardware is added to an existing camera, which would require updates to the API to allow use of the new features. However, after analysing the changes and from the information gathered in the interview, we discovered that hardware has played a minimal role in the evolution of the API.

@@

\noindent
\textbf{Functionality is reverse engineerable  } %Reviewed by JW
All changes whose main motive was to either add or remove functionality were accurately predicted. These changes were the most significant and it was expressed during the interview that these types of changes were the main driving factors behind the development of the new version. In summary, eleven out of a total of thirteen changes were correctly predicted. It is also interesting to note that all changes which were not assigned any sub-category were correctly predicted. This supports the fact that such changes have a clearer intent and are perhaps more well-defined.

\smallskip \noindent
\textbf{Design changes are reverse engineerable to some extent  }  %Reviewed by JW
Three out of five changes which did not add functionality, but rather were introduced to reduce complexity and increase quality, were correctly predicted. The two changes which were incorrectly predicted were inaccurate because the main motivation was confused with the secondary intent. This shows that, although the prediction was incorrect, the general reasoning was correct. Our predictions were accurate enough to allow us to form a reliable conclusion which provided valuable discussion with the case company.

@@

\noindent
\textbf{A majority of the laws apply  } %Reviewed by JW
Sufficient evidence was found for the laws of Continuing Change, Self Regulation, Conservation of Familiarity, Continuing Growth and Feedback System to establish that they apply to APIs. The most prominent explanation for this is that a large part of the introduced changes were related to functionality. It is especially interesting to note that the laws of Self Regulation and Conservation of Familiarity accurately match the API explored in this paper.

\smallskip \noindent
\textbf{Two laws apply to a certain extent  } %Reviewed by JW
The laws of Increasing Complexity and Declining Quality were found to apply to APIs, but the extent to which they apply might not be as significant for APIs compared to other types of systems. The reason that this doubt exists is because these motives were secondary to the motives of increased functionality. An alternative explanation for these laws being of lesser significance in APIs could be that the general inertia to modify APIs which exists \cite{google_talk} \cite{henning2007api} \cite{mcdonnell2013empirical} \cite{robbes2012developers} might dominate the need to improve simplicity and quality.

\smallskip \noindent
\textbf{One law is not applicable  } %Reviewed by JW
Strong evidence was found to suggest that the law of Conservation of Organisational Stability does not apply to APIs. Our results not only show that the law is not true for APIs, but that the opposite of what the law states is true for APIs. This is supported by the consensus in literature of the negative effects caused by modifications to APIs \cite{google_talk} \cite{henning2007api} \cite{mcdonnell2013empirical} \cite{robbes2012developers}. The results from our study also revealed that reluctance to change APIs causes the releases of new versions to be few and sporadic.

\smallskip \noindent
\textbf{The importance of a Feedback System  } %Reviewed by JW
Especially interesting results were found in regards to the law of Feedback System. The importance of having a structured and developed feedback loop was stressed during the interview and it was also expressed as one of the main points of improvement. This suggests that including feedback from more sources and using this feedback to improve the evolution process should be a consideration for APIs.
